---
title: "621 final"
author: "Tyler Baker and Jay Lee"
date: '2022-11-20'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(psych)
library(ggplot2)
library(GGally)
library(pscl)
library(reshape)
library(car)
library(mice)
library(naniar)
library(readr)
library(dplyr)
library(knitr)
library(tidyr)
library(stringr)
library(extraoperators)
library(palmerpenguins)
```

```{r}
ins_df <- 'https://raw.githubusercontent.com/jayleecunysps/sps621/main/insurance.csv'
ins_df <-read.csv(ins_df, header=T, stringsAsFactors=T)
```

We want to check the type each columns.

age: age of primary beneficiary (int)

sex: insurance contractor gender, female, male (Factor)

bmi: Body mass index, providing an understanding of body, weights that are relatively high or low relative to height,
objective index of body weight (kg / m ^ 2) using the ratio of height to weight, ideally 18.5 to 24.9 (num)

children: Number of children covered by health insurance / Number of dependents (int)

smoker: Smoking (Factor)

region: the beneficiary's residential area in the US, northeast, southeast, southwest, northwest. (Factor)

charges: Individual medical costs billed by health insurance (num)
```{r}
dim(ins_df) #check how many obs and var
str(ins_df) #check the data type
```

In the summary, we can see the range of charges and bmi is huge, also the amount of smoker and non-smoker has huge gap in between.

```{r}
summary(ins_df)
colSums(is.na(ins_df)) #check if there is any missing value
```
As we can see, there is a lot of outliner under charges and bmi which support what we saw under summary.

```{r}
boxplot(ins_df$age,
data=ins_df,
main=paste('Box Plot - age'),
xlab='age')
boxplot(ins_df$bmi,
data=ins_df,
main=paste('Box Plot - bmi'),
xlab='bmi')
boxplot(ins_df$children,
data=ins_df,
main=paste('Box Plot - children'),
xlab='children')
boxplot(ins_df$charges,
data=ins_df,
main=paste('Box Plot - charges'),
xlab='charges')
```
As we can see most of the people charges are under 20k.
```{r}
ggplot(data = ins_df, aes(x = charges)) + 
  geom_density() + 
  ggtitle("Distribution of Charges")
```
sex and region do not have big differences while smoke drive the differences a lot.

```{r}
ggplot(data=ins_df, aes(x=sex, y=charges, fill=sex)) +
geom_bar(stat="identity", position=position_dodge())
ggplot(data=ins_df, aes(x=smoker, y=charges, fill=sex)) +
geom_bar(stat="identity", position=position_dodge())
ggplot(data=ins_df, aes(x=region, y=charges, fill=sex)) +
geom_bar(stat="identity", position=position_dodge())
```
We would like the see the distribution again with "fill = smoker" since we notied see the noticeable differences.
Now we can see smoker pay much more charges then non-smoker.

```{r}
ggplot(data = ins_df, aes(x = charges,fill = smoker)) + 
  geom_density() + 
  ggtitle("Distribution of Charges")
```

then we want to see scatterplot with fitted regression line, the bmi one catch our attention.

```{r}
ggplot(ins_df, aes(x = age, y = charges, fill=smoker)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
ggplot(ins_df, aes(x = bmi, y = charges, fill=smoker)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
ggplot(ins_df, aes(x = children, y = charges, fill=smoker)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```

```{r}
model1 <- lm(ins_df$charges~ins_df$bmi)
summary(model1)
plot(ins_df$bmi, ins_df$charges, main = "Height vs Weight
                 Regression model")
abline(lm(ins_df$charges~ins_df$bmi))
```

```{r}
model2 <- lm(charges~bmi + age +children +region,
            data = ins_df)
summary(model2)
summary(model2)$coefficient
```

the t value of children and region is closest to zero, so we decide to elimit children and create model3.

model equation is charges = -6424.80 + 332.97 * bmi + 241.93 * age
```{r}
model3 <- lm(charges~bmi + age +children,
            data = ins_df)
summary(model3)
summary(model3)$coefficient
```

```{r}
model3v2 <- lm(charges~bmi + age,
            data = ins_df)
summary(model3v2)
summary(model3v2)$coefficient
```



```{r}
set.seed((1212))

partition <- sample(1:nrow(ins_df), size=nrow(ins_df)*0.7,replace=FALSE)

train.data <-ins_df[partition, ]
test.data <- ins_df[-partition, ]

target <- test.data$charges
test.data <- test.data[, c("age", "sex", "bmi", "children", "smoker", "region")]

model4 <- glm(charges ~.,family = gaussian(link="identity"),data=train.data)

predictions <- as.data.frame(predict(model4, newdata = test.data))
accuracy_df <- target

accuracy_df<- merge(accuracy_df, predictions)

accuracy_df$error <- abs(accuracy_df$x - accuracy_df$`predict(model4, newdata = test.data)`)
accuracy_df$percentage_error <- ((accuracy_df$`predict(model4, newdata = test.data)` - accuracy_df$x)/accuracy_df$x) * 100
avg_error <- mean(accuracy_df$error)
avg_percentage_error <- mean(accuracy_df$percentage_error)
print(avg_percentage_error)
print(avg_error)
summary(model4)
```



